---
title: "Introducing `alarmdata`"
description: |
  The first stable version of our package `alarmdata` is now on CRAN, introducing
  a data-focused package for using the outputs of ALARM Project research.
author:
    - name: Cory McCartan
      affiliation: Center for Data Science, New York University
      location: Cambridge, MA
      email: corymccartan@nyu.edu
      url: https://www.corymccartan.com
    - name: Christopher T. Kenny
      affiliation: Department of Government, Harvard University
      location: Cambridge, MA
      email: christopherkenny@fas.harvard.edu
      url: https://www.christophertkenny.com/
    - name: Tyler Simko
      affiliation: Department of Government, Harvard University
      location: Cambridge, MA
      email: tsimko@g.harvard.edu
      url: https://tylersimko.com/
    - name: Michael Zhao
      affiliation: Harvard College
      location: Cambridge, MA
      email: michaelzhao@college.harvard.edu
    - name: Kosuke Imai
      affiliation: Departments of Government and Statistics, Harvard University
      location: Cambridge, MA
      email: imai@harvard.edu
      url: https://imai.fas.harvard.edu/
date: 2024-03-06
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

`alarmdata` is the newest component of the [`redistverse`](https://alarm-redist.org/redistverse), a family of packages which provide tools for redistricting analyses in R. 
As of the past weekend, it is now on CRAN.
You can install the following with:

```r
install.packages('alarmdata')
```


# Downloading data

`alarmdata` provides several functions for loading in data from ALARM Project projects. 
Most useful are `alarm_50state_map()` and `alarm_50state_plans()` which respectively load in the maps and plans from our [50-states project](https://alarm-redist.org/fifty-states/).

To download and use these data, you can use the following code:

```{r}
library(alarmdata)
library(redist)

map_wa <- alarm_50state_map('WA')
plans_wa <- alarm_50state_plans('WA')
```


# Using `redist_plans` with your own data

Once you've downloaded some plans, we've made it easy for you to add your own data to the `redist` framework.
To demo this, we can look at the NY plans passed last week (February 2024).

First, we can load in the map and plans for New York:

```{r}
map_ny <- alarm_50state_map('NY')
plans_ny <- alarm_50state_plans('NY')
```

Next, we can read in the block assignment data for the commission plan and the legislature plan.
For convenience of the example, we can grab this data from a [GitHub repo](https://github.com/christopherkenny/ny-baf). 

```{r}
# commission link
comm_link <- 'https://raw.githubusercontent.com/christopherkenny/ny-baf/main/data/A09310.csv'

# download the legislature xlsx file
leg_link <- 'https://github.com/christopherkenny/ny-baf/raw/main/data-raw/congressional_plan_equivalency.xlsx'
temp_leg <- tempfile(fileext = '.xlsx')
curl::curl_download(leg_link, temp_leg)

# read in the data
baf_commission <- readr::read_csv(comm_link, col_types = 'ci') |> 
  dplyr::rename(A09310 = district)
baf_legislature <- readxl::read_excel(temp_leg, col_types = c('text', 'text')) |> 
  dplyr::rename(
    GEOID = Block,
    commission2024 = `DistrictID:1`
  ) |> 
  dplyr::mutate(commission2024 = as.integer(commission2024))
```

Finally, `alarmdata` provides a function `alarmdata::alarm_add_plan()` to add the new reference plans to the underlying `redist_plans` object.

```{r}
plans_ny <- plans_ny |>
  alarm_add_plan(ref_plan = baf_commission, map = map_ny, name = 'A09310') |>
  alarm_add_plan(ref_plan = baf_legislature, map = map_ny, name = 'commission2024')
```

This gives a full `redist_plans` object with the new plans added in.
We can do anything we want with this.
For example, stealing a bit of code from [](), we can look at the Democratic vote shares across districts:

First, we can define some helper functions for the plot:

```{r}
lbl_party <- function(x) {
  dplyr::if_else(x == 0.5, "Even",
    paste0(dplyr::if_else(x < 0.5, "R+", "D+"), scales::number(200 * abs(x - 0.5), 1))
  )
}

r_geom <- function(...) {
  ggplot2::geom_point(
    ggplot2::aes(x = as.integer(.data$.distr_no),
                 y = e_dvs,
                 color = .data$draw, 
                 shape = .data$draw),
    ...
  )
}
```

Then, we can make a simple plot of Democratic vote share by district:

```{r}
library(ggplot2)
library(ggredist)
redist.plot.distr_qtys(plans_ny, e_dvs,
                       color_thresh = 0.5,
                       size = 0.04 - sqrt(8) / 250, alpha = 0.4,
                       ref_geom = r_geom
) +
    geom_hline(yintercept = 0.5, color = '#00000055', size = 0.5) +
    scale_y_continuous('Two-party vote margin', labels = lbl_party) +
    labs(x = 'Simulated districts, ordered by Democratic vote margin') +
    annotate('text',
             x = 3.5, y = sort(plans_ny$e_dvs[1:26])[3],
             label = 'Commission 2024', hjust = -0.05, size = 3.5,
             color = '#52796F'
    ) +
    annotate('text',
             x = 1.5, y = min(plans_ny$e_dvs[27:52]),
             label = 'A09310', hjust = 0.05, size = 3.5,
             color = '#A09310'
    ) +
    annotate('text',
             x = 3.5, y = sort(plans_ny$e_dvs[53:78])[3],
             label = 'Cervas', hjust = -0.05, size = 3.5,
             color = 'black'
    ) + 

    scale_color_manual(values = c('#52796F', '#A09310', 'black', ggredist$partisan[2], ggredist$partisan[14]),
                       labels = c('pt', 'Rep.', 'Dem.'), guide = 'none') +
    scale_shape_manual(values = c(16, 17, 18), guide = 'none') +
    theme_bw()
```

# Cache data across projects

By default, all downloads are directed to a temporary directory.
Each time you reload R, you need to re-download the data.
To avoid this, you can set the `alarm.use_cache` option to a directory where you want to store the data.

```r
options(alarm.use_cache = TRUE)
```

This can be set on a by-session basis, but we recommend setting it in your `.Rprofile` so that it is set every time you start R. To open this file, use the following command:

```r
usethis::edit_r_profile()
```
