---
title: "redist 101"
author: "Sho Miyazaki, Kevin Wang, Melissa Wu, Kento Yamada"
format: html
execute:
  freeze: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

<br> *"There are a handful of college professors who even have the expertise to run [redistricting] simulations in the first place."* <br>

::: {align="right"}
---lawyer from oral arguments before the Supreme Court in [*Merrill v. Milligan*](https://www.scotusblog.com/case-files/cases/merrill-v-milligan-2/)
:::

<br>

```{r echo = FALSE}
knitr::include_graphics("https://alarm-redist.org/fifty-states/AL_cd_2020/index_files/figure-html/enacted-samples-1.svg")
```

This vignette introduces algorithmic redistricting analysis using the `redist` package. It also demonstrates the `alarmdata` package, which downloads ready-to-use redistricting data and simulations from the ALARM Project's [50-State Redistricting Simulations](https://alarm-redist.org/fifty-states/).

## Overview

### Why Simulate?

When attempting to detect gerrymandering, we want to determine whether a certain district map is unfair. But what makes a map unfair, and how do we identify whether a map is unfair? The problem is that it is not obvious what standard to which we should compare an existing map.

The key challenge is constructing a counterfactual. We want to compare an existing map to a set of alternative redistricting plans that are drawn following the same rules as those used in the actual redistricting process.

One potential solution is to list all possible plans. Unfortunately, this is impossible in most cases because the number of potential plans tends to be astronomically large. Even under a simple scenario in which an 8-by-8 checkerboard is split into two contiguous districts, the number of unique plans totals over $1.2 \times 10^{11}$ (Fifield et al., 2020). In real-world redistricting problems, we often deal with thousands of administrative units. Thus, enumerating all potential plans is unrealistic.

Another idea might be to compare the plan under question to plans adopted in other states, or to plans that were used in the same state in the past. This is, again, is a misleading approach because the geographical distributions of voters differ across different states and different time periods. Moreover, each state follows its own rules for redistricting, and we cannot justify comparing plans that were drawn under different rules (Kenny et al., 2022).

Therefore, we would want to sample plans that follow the same rules used in the real redistricting process and that are based on the real distribution of voters in the same state, in the same time period.

Simulations make this possible. Using simulations, we can obtain a representative sample of redistricting plans that follow the constraints that are used in the actual redistricting process such as:

-   Contiguity (Making sure that districts are contiguous)
-   Population parity (Making sure that districts have equal population)
-   Limiting administrative boundary splits
-   Other idiosyncratic requirements, such as connecting counties by transportation links

Typically, we draw a few thousand plans in each state, making sure that the plans are unique and that the sample is representative. Then, we compare the simulated plans with the enacted/proposed map to see whether, for example, the map under question disproportionately favors one group over another. For a more intuitive explanation of simulations, check out this Washington Post article titled ["Can computer simulations help fix democracy?"](https://www.washingtonpost.com/politics/interactive/2022/algorithmic-redistricting/). The article talks about how Prof. Kosuke Imai, head of the ALARM Project, served as an expert witness in a redistricting litigation in Ohio.

### Evaluating Plans

After we obtain a representative sample of redistricting plans, we can detect whether the map under question was gerrymandered by comparing the map with the simulated plans. `redist` and its associated packages are equipped with tools that allow you to conduct and visualize such analysis. The metrics used in such analysis include:

-   The number of seats that each party is expected to win under the enacted/proposed plan and the simulated plans
-   The two-party vote margin in each district, which is useful when trying to identify whether specific groups of voters have been packed into a small number of districts or splintered into numerous districts
-   The efficiency gap, which represents the degree to which the votes cast for each party are wasted
-   The partisan bias of plans, which quantifies the excess seat share per party relative to the party's vote share

We can conclude that there is evidence of gerrymandering if the map under question is a clear outlier relative to the simulated plans. Check out how we conduct these types of analysis in our [50-State Simulation Project](https://alarm-redist.org/fifty-states/about/).

### References and Further Reading

Fifield, B., Imai, K., Kawahara, J., & Kenny, C. T. (2020). The Essential Role of Empirical Validation in Legislative Redistricting Simulation. *Statistics and Public Policy, 7*(1), 52--68. <https://doi.org/10.1080/2330443X.2020.1791773>

Kenny, C. T., McCartan, C., Simko, T., Kuriwaki, S., & Imai, K. (2022). Widespread Partisan Gerrymandering Mostly Cancels Nationally, but Reduces Electoral Competition. <https://doi.org/10.48550/ARXIV.2208.06968>

## Redistricting Simulation Workflow

Before starting, we need to load the following packages, which contain the functions necessary to conduct the simulations.

```{r}
library(tidyverse)
library(alarmdata)
library(redist)
```

## Building Data Objects

### `redist_map`

The first step in a `redist` analysis is to create the `redist_map` object, which stores the basic parameters of the redistricting problem. Here, we use `redist_map` objects that can be downloaded from the [50-State Redistricting Simulations](https://alarm-redist.org/fifty-states/) using the `alarmdata` package.

```{r message=FALSE}
map_nc <- alarm_50state_map("NC")
print(map_nc)
```

As shown above, `redist_map` (and shapefiles) looks like a dataframe with many columns, such as population and geographical information.

For example, the `geometry` column contains the geographic shapefile information, which can be mapped as a `POLYGON` (or `MULTIPOLYGON`) shape of the geographical area.

```{r}
head(map_nc$geometry)
```

### Shapefiles and Projections

If you want to use the other shapefile data for the redistricting analysis (other countries, etc.), you need to transform the available shapefiles (census data, etc.) into a `redist_map` object.

As there are multiple formats of [shapefiles](https://r-spatial.github.io/sf/), you may also need to convert and adjust them into a unified format. For example, projections must be in the same style. For more information, check pp.17-18 of [this presentation for the workshop](https://docs.google.com/presentation/d/1FlTfDv8sjhDxxU2KIOnKd1SHhvivm7WdD0bYdrNNgo0/edit#slide=id.g11d65f9e588_0_0).

### Adjacency List

Since districts are generally required to be contiguous, all `redist` algorithms require an *adjacency graph*, which is constructed from the geography of the actual precincts (or other redistricting unit).

What is adjacency? It is the formalization of contiguity. Two units are adjacent if they are contiguous geographically.

Let's check what the adjacency list looks like. The `adj` column of `redist_map` is the adjacency list.

```{r}
head(map_nc$adj)
```

As you can see, the adjacency list of each precinct contains all other precincts to which it is adjacent/with which it is contiguous.

Using the `redist_map` object, we can plot this adjacency list.

```{r}
plot(map_nc, adj = TRUE)
```

You may need to fix the adjacency list manually with the [`geomander` package](https://christophertkenny.com/geomander/reference/add_edge.html), particularly when there are bodies of water or islands connected by bridges.

### Further Pre-Processing

Often, we want to only analyze a portion of a map, or hold some districts fixed while others are re-simulated. We may also want to implement a status-quo-type constraint that encourages simulated districts to be close to a reference plan. This can be accomplished by freezing the "cores" of each district.

All of these operations fall under the umbrella of map pre-processing, and `redist` is well-equipped to handle them.

The map [pre-processing vignette](https://alarm-redist.org/redist/articles/map-preproc.html) contains more information and examples about these operations.

## Simulating redistricting plans

### Sampling Algorithms

As discussed above, we want to create counterfactual plans by obtaining a representative sample with simulation algorithms. The `redist` package has three functions that use Sequential Monte Simulations (`redist_smc`) and Markov Chain Monte Carlo (`redist_flip` and `redist_mergesplit`) algorithms. In this section, we will pick up `redist_smc`, which generates nearly independent congressional or legislative redistricting plans according to contiguity, population, compactness, and administrative boundary constraints. For the details of the functions, check the following.

-   [`redist_smc`](https://alarm-redist.org/redist/reference/redist_smc.html)
-   [`redist_flip`](https://alarm-redist.org/redist/reference/redist_flip.html) and \*[`redist_mergesplit`](https://alarm-redist.org/redist/reference/redist_mergesplit.html)

### `redist_smc`

Once the `redist_map` is ready, we can run the simple simulations. Here, let's try generating five hundred simulated plans for North Carolina. Note that we can ask `redist_smc` to produce five hundred plans by combining two separate simulation runs. This ensures that the algorithm "converges"--that is, that the distribution of plans we obtain is stable.

```{r}
smc_plans_basic <- redist_smc(map_nc, nsims = 500, runs = 4L) |>
    mutate(
        plan_dev = plan_parity(map_nc),
        comp_edge = distr_compactness(map_nc),
        county_splits = county_splits(map_nc, map_nc$county),
        e_dvs = partisan_metrics(map_nc, "DVS", nrv, ndv)
    )
```

Now, you have got 2,000 plans. The output of the simulations is `redist_plans` format. The followings are the first six plans generated by the simulations, and you can plot them.

```{r}
redist.plot.plans(smc_plans_basic, draws = 1:6, shp = map_nc)
```

Using the `summary` function, we can perform a diagnostic check to ensure that our plans have converged and are diverse (in other words, that we are exploring a sufficient space of possible alternative plans, rather than simulating the same plan over and over again).

```{r}
summary(smc_plans_basic)
```

### Constraints

The simulation output from the SMC is unbiased random sample, but not always mean representative. Since each state (or equivalent) has a different legal requirements for the redistricting plan, representative sample has to be in the resemble the space of legal possibilities. Thus, we need to customize the constraints to operationalize legal requirements and norms. For example, in some cases, each district has to have an equal population, while the margins of tolerance exist, which can be varied by the state. To adjust to that rule, we can customize the constraints as you can see below. In `redist`, you can add both *hard* and *soft* constraints.

#### Hard Constraints

-   Population Tolerance (`pop_tol`= 0.0x) -- all plans must be within x% of population equality.
-   Contiguity -- all plans must be contiguous
-   County splits -- all plans must have up to a small number of splits

#### Soft Constraints

-   `redist` allows you to add soft constraints that are listed [here](https://alarm-redist.org/redist/reference/constraints.html)
-   Using [`add_constr_xyz()`](https://alarm-redist.org/redist/reference/constraints.html) (`xyz` varies depending on the constraints you want to add) functions, you can add constraints on the redistricting process to be encoded in the target distribution for sampling of `redist_smc()` and `redist_mergesplit()`
-   Note that higher and higher strength values will eventually cause the algorithm's accuracy and efficiency to suffer, so be sure to check all sampling diagnostics.

### Example

#### Change Population Tolerance

You can use `get_pop_tol()` function to check the current population tolerance of `redist_map` object.

```{r}
get_pop_tol(map_nc)
```

As you can see above, the population tolerance of `map_nc` is `r get_pop_tol(map_nc)`. This means that all plans must be within `r 100*get_pop_tol(map_nc)`% of population equality. This is why you can see the output message of `redist_smc()` shows following.

```{r}
redist_smc(map_nc, 10)
```

As you can see, all the `total_pop` should be within between 741,942 and 749,399, which is `r 100*get_pop_tol(map_nc)`% of population equality.

Then, let's change the population tolerance with `set_pop_tol()` function. Here, we will set `pop_tol` into 0.05 (5%).

```{r}
map_nc_constr <- set_pop_tol(map_nc, 0.05)
redist_smc(map_nc_constr, 10)
```

Now, as the population tolerance changes, the possible range of `total_pop` expanded between 708,387 and 782,954.

#### Administrative Boundary Splits

We can also add a soft constraints to the simulation. Here, we will try to add a constraint that prefers the fewer county splits by creating `redist_constr` object.

```{r}
constr <- redist_constr(map_nc_constr)
constr <- add_constr_splits(constr, strength = 1.5, admin = county)
smc_plans_constr <- redist_smc(
    map = map_nc,
    nsims = 500,
    runs = 4L,
    constraints = constr
)
```

### `redist_plans` Output

Changing the constraints means changing the output sample of the simulations. Although we are going to explain in detail how to evaluate and analyze the simulation output with certain measures, here let's see how the constraints changed the population deviation of the simulation outcome.

```{r fig.show="hold", out.width="50%", fig.align='center'}
# original constraints
smc_plans_basic |> 
    group_by(draw) |>
    summarize(pop_dev = max(abs(total_pop / mean(total_pop) - 1))) |>
    redist.plot.hist(pop_dev, bins = 10) +
    xlim(c(0.003, 0.0055)) +
    theme_bw()

# customized constraints
smc_plans_constr |> 
    group_by(draw) |>
    summarize(pop_dev = max(abs(total_pop / mean(total_pop) - 1))) |>
    redist.plot.hist(pop_dev, bins = 10) +
    xlim(c(0.003, 0.0055)) +
    theme_bw()
```

## North Carolina Analysis Example

We will now demonstrate an example of redistricting analysis using the simulated plans from the 50-States Project. 

### Setup

First, we load the required packages for analysis. We also load the code for generating the validation plots, which will help us create visualizations for the summary statistics of the simulations.

```{r, include=TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(alarmdata)
library(redist)
library(geomander)
library(here)
library(sf)

source("fifty-states/validate.R")
```

### Download Simulation Data

We can access data from all 50 states in the US using the `alarmdata` package. Here, we use North Carolina as an example: `alarm_census_vest()` downloads geographic, demographic, and election data, `alarm_50state_map()` loads the `redist_map` object that contains precinct geometries and adjacencies, and `alarm_50state_plans()` downloads the `redist_plans` object that contains 5000 pre-simulated redistricting plans from the [50-State Redistricting Simulations](https://doi.org/10.48550/arXiv.2206.10763).

```{r}
data_nc <- alarm_census_vest("NC")
map_nc <- alarm_50state_map("NC")
plans_nc <- alarm_50state_plans("NC")
```

The code for how these plans were generated is available on the [50-State GitHub repository](https://github.com/alarm-redist/fifty-states/tree/main/analyses/NC_cd_2020).

Specifically in North Carolina, districts must:
1) be contiguous
2) have equal populations
3) be geographically compact
4) preserve county boundaries as much as possible.
Each of these redistricting requirements, in addition to compliance with the 1965 Voting Rights Act, were closely followed when conducting the simulations.

### Visualization

We can plot some of the simulated plans to visualize the geographical boundaries of the simulated districts.

```{r}
redist.plot.plans(plans_nc, draws = 1:4, shp = map_nc)
```

We can also plot the VI distance of each plan, which measures how different the simulated plans are from one another. In the following plot, the VI distances are mostly within the range 0.5-1.0, which means that the simulated plans include a diverse set of geographical arrangements.

```{r}
plan_div <- plans_diversity(plans_nc, n_max = 150)
qplot(plan_div, bins = I(40), xlab = "VI distance", main = "Plan diversity") + 
    theme_bw()
```

We can also visualize the compactness of entire plans. The following plot shows the compactness of the simulated plans at the plan level, using the Fraction of Edges Kept measure. The compactness of the NC 2020 enacted plan is within the mid-upper range of the simulated plans, suggesting that the plans were constructed with appropriate compactness considerations.

```{r}
redist.plot.hist(plans_nc, qty = comp_edge) +
    labs(x = "Fraction of Edges Kept", y = "Percentage of Plans") +
    theme_bw()
```

The following plot displays the distribution of compactness across the simulated plans at the district level, using the Polsby-Popper compactness measure.

```{r}
redist.plot.distr_qtys(plans_nc, qty = comp_polsby, geom = "boxplot") +
    labs(y = "Compactness: Polsby-Popper") +
    theme_bw()
```

### Simulation Evaluation

To evaluate the performance of the simulations, we can use `summary(plans)` to check for any bottlenecks or efficiency losses. Be sure to check that all R-hat values are below 1.05 and that the SMC runs have converged. For NC, 2 runs of 10,000 simulations is the typical size needed for convergence. Example summary statistics for fully converged simulation runs are pasted below.

```{r}
#| eval: false
summary(plans_nc)
```

```
SMC: 20,000 sampled plans of 14 districts on 2,666 units
`adapt_k_thresh`=0.985 • `seq_alpha`=0.5
`est_label_mult`=1 • `pop_temper`=0.01

Plan diversity 80% range: 0.69 to 0.87

R-hat values for summary statistics:
pop_overlap      total_vap       plan_dev      comp_edge    comp_polsby       pop_hisp      pop_white      pop_black       pop_aian 
1.011740       1.004010       1.016022       1.006402       1.007232       1.011292       1.001015       1.008898       1.000115 
pop_asian       pop_nhpi      pop_other        pop_two       vap_hisp      vap_white      vap_black       vap_aian      vap_asian 
1.027659       1.001441       1.040853       1.000344       1.011211       1.001468       1.005998       1.000307       1.030569 
vap_nhpi      vap_other        vap_two pre_16_rep_tru pre_16_dem_cli uss_16_rep_bur uss_16_dem_ros gov_16_rep_mcc gov_16_dem_coo 
1.006742       1.039742       1.003739       1.014649       1.041666       1.011259       1.037633       1.011370       1.036032 
atg_16_rep_new atg_16_dem_ste sos_16_rep_lap sos_16_dem_mar pre_20_rep_tru pre_20_dem_bid uss_20_rep_til uss_20_dem_cun gov_20_rep_for 
1.012634       1.036626       1.010377       1.033683       1.012790       1.048726       1.009076       1.047310       1.013620 
gov_20_dem_coo atg_20_rep_one atg_20_dem_ste sos_20_rep_syk sos_20_dem_mar         arv_16         adv_16         arv_20         adv_20 
1.046216       1.009898       1.044914       1.009182       1.044455       1.010635       1.036324       1.010868       1.046275 
county_splits    muni_splits            ndv            nrv        ndshare          e_dvs         pr_dem          e_dem          pbias 
1.004991       1.014032       1.043048       1.012762       1.028974       1.028886       1.044242       1.006234       1.006050 
egap 
1.004837 

Sampling diagnostics for SMC run 1 of 2
Eff. samples (%) Acc. rate Log wgt. sd  Max. unique Est. k 
Split 1     8,041 (80.4%)     11.5%        0.43 6,354 (101%)     12 
Split 2     7,829 (78.3%)     17.7%        0.54 5,974 ( 95%)      7 
Split 3     7,698 (77.0%)     22.3%        0.61 5,852 ( 93%)      5 
Split 4     7,560 (75.6%)     31.0%        0.65 5,867 ( 93%)      3 
Split 5     7,423 (74.2%)     18.4%        0.68 5,671 ( 90%)      5 
Split 6     7,445 (74.5%)     20.7%        0.70 5,740 ( 91%)      4 
Split 7     7,552 (75.5%)     19.2%        0.72 5,709 ( 90%)      4 
Split 8     7,442 (74.4%)     23.1%        0.71 5,675 ( 90%)      3 
Split 9     7,390 (73.9%)     21.1%        0.73 5,696 ( 90%)      3 
Split 10    7,515 (75.2%)     14.3%        0.73 5,664 ( 90%)      4 
Split 11    7,669 (76.7%)     15.5%        0.72 5,482 ( 87%)      3 
Split 12    7,820 (78.2%)     11.8%        0.70 5,334 ( 84%)      3 
Split 13    7,423 (74.2%)      4.0%        0.77 4,850 ( 77%)      3 
Resample    2,891 (28.9%)       NA%        1.21 4,691 ( 74%)     NA 

Sampling diagnostics for SMC run 2 of 2
Eff. samples (%) Acc. rate Log wgt. sd  Max. unique Est. k 
Split 1     8,087 (80.9%)     15.5%        0.43 6,267 ( 99%)      9 
Split 2     7,872 (78.7%)     20.2%        0.54 5,938 ( 94%)      6 
Split 3     7,650 (76.5%)     21.9%        0.60 5,908 ( 93%)      5 
Split 4     7,565 (75.7%)     24.5%        0.65 5,812 ( 92%)      4 
Split 5     7,380 (73.8%)     18.5%        0.68 5,786 ( 92%)      5 
Split 6     7,431 (74.3%)     27.1%        0.71 5,693 ( 90%)      3 
Split 7     7,455 (74.6%)     19.4%        0.73 5,683 ( 90%)      4 
Split 8     7,509 (75.1%)     23.0%        0.73 5,720 ( 90%)      3 
Split 9     7,523 (75.2%)     12.8%        0.71 5,734 ( 91%)      5 
Split 10    7,509 (75.1%)     13.9%        0.70 5,664 ( 90%)      4 
Split 11    7,403 (74.0%)      9.4%        0.74 5,542 ( 88%)      5 
Split 12    7,589 (75.9%)     11.9%        0.73 5,351 ( 85%)      3 
Split 13    7,744 (77.4%)      4.1%        0.72 4,926 ( 78%)      3 
Resample    3,593 (35.9%)       NA%        1.15 4,889 ( 77%)     NA
```

We can also run `validate_analysis(plans, map)` to generate a set of useful simulation visualizations.

```{r fig.height = 12, fig.width = 10}
validate_analysis(plans_nc, map_nc)
```

### Comparing Custom Plans

The 50-State `redist_plans` objects include data for the 2020 enacted districting plans. However, analysts might want to compare other custom plans to the 50-State simulations. This can be done with `alarm_add_plan()`. The following example adds one of the previously proposed 2020 congressional maps for North Carolina as a reference plan to the `plans_nc` object. Many submitted maps are publicly available online on sites such as [Redistricting Data Hub](https://redistrictingdatahub.org/data/download-data/#state-menu), [All About Redistricting](https://redistricting.lls.edu/mapdownload/), or state government websites. The North Carolina enacted and proposed maps came from the [North Carolina General Assembly](https://www.ncleg.gov/Redistricting).

```{r}
# Download plan from NC General Assembly
fs::dir_create("plans")
path_enacted <- here("plans", "SL 2021-174 Congress.shp")
if (!file.exists(path_enacted)) {
    url <- "https://s3.amazonaws.com/dl.ncsbe.gov/ShapeFiles/USCongress/2021-11-04%20US_Congress_SL_2021-174.zip"
    download.file(url, paste0(dirname(path_enacted), "/nc.zip"))
    unzip(paste0(dirname(path_enacted), "/nc.zip"), exdir = dirname(path_enacted))
}

# Add plan to redist_map
dists <- read_sf(path_enacted)
dists <- st_transform(dists, st_crs(map_nc))
map_nc$cd_2020_ex <- as.integer(dists$DISTRICT)[geo_match(from = map_nc, to = dists, method = "area")]

# Add plan as reference using `alarm_add_plan()`
plans_nc <- plans_nc |>
    alarm_add_plan(ref_plan = map_nc$cd_2020_ex, map_nc, name = "cd_2020_ex")
```

After the old reference plan has been added, we can compare the simulation statistics to both the previously proposed plan, labeled `cd_2020_ex`, and the currently enacted plan, labeled `cd_2020`. The following code plots the number of county splits, and we can see that the current plan actually splits more counties than the previous plan.

```{r}
hist(plans_nc, county_splits) + 
    labs(title = "County splits") + 
    theme_bw()
```

We can compare distributions of different summary statistics across districts. The following boxplots display the distribution of Democratic vote percentage across each district. This plot shows that there are more majority-Democratic districts in the current plan than in the previously ratified plan, due to less packing of voters.

```{r}
redist.plot.distr_qtys(
    plans_nc, 
    qty = pre_20_dem_bid / (pre_20_dem_bid + pre_20_rep_tru), 
    geom = "boxplot"
) +
    labs(y = "2020 Presidential Democratic Vote Percentage") +
    theme_bw()
```

Similarly, we can visualize the minority voting age population (VAP) percentage in each district. This plot shows that voters are less packed in the highest MVAP districts, leading to more opportunity districts.

```{r}
redist.plot.distr_qtys(
    plans_nc, 
    qty = (total_vap - vap_white) / (total_vap), 
    geom = "jitter"
) +
    labs(y = "2020 Minority VAP Percentage") +
    theme_bw()
```

Overall, although the newly enacted plan split more counties than the previously proposed plan, the partisan and demographic metrics of the current plan are not among the far outliers of the simulation diagnostics, suggesting that the current plan is a more representative configuration given the specific redistricting requirements in North Carolina.
