---
title: "Introducing `alarmdata`"
description: |
  The first stable version of our package `alarmdata` is now on CRAN, introducing
  a data-focused package for using the outputs of ALARM Project research.
author:
    - name: Cory McCartan
      affiliation: Center for Data Science, New York University
      location: Cambridge, MA
      email: corymccartan@nyu.edu
      url: https://www.corymccartan.com
    - name: Christopher T. Kenny
      affiliation: Department of Government, Harvard University
      location: Cambridge, MA
      email: christopherkenny@fas.harvard.edu
      url: https://www.christophertkenny.com/
    - name: Tyler Simko
      affiliation: Department of Government, Harvard University
      location: Cambridge, MA
      email: tsimko@g.harvard.edu
      url: https://tylersimko.com/
    - name: Michael Zhao
      affiliation: Harvard College
      location: Cambridge, MA
      email: michaelzhao@college.harvard.edu
    - name: Kosuke Imai
      affiliation: Departments of Government and Statistics, Harvard University
      location: Cambridge, MA
      email: imai@harvard.edu
      url: https://imai.fas.harvard.edu/
date: 2024-03-10
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`alarmdata` is the newest component of the [`redistverse`](https://alarm-redist.org/redistverse), a family of packages which provide tools for redistricting analyses in R. 
As of the past weekend, it is now on CRAN.
You can install the following with:

```r
install.packages('alarmdata')
```


# Downloading data

`alarmdata` provides several functions for loading in data from ALARM Project projects. 
Most useful are `alarm_50state_map()` and `alarm_50state_plans()` which respectively load in the maps and plans from our [50-states project](https://alarm-redist.org/fifty-states/).
The map object contains Census and election for each state precinct.
The plans object contains alternative simulated plans.

To download and use these data, you can use the following code:

```{r}
library(alarmdata)
library(redist)

map_wa <- alarm_50state_map('WA')
plans_wa <- alarm_50state_plans('WA')
```


# Using `redist_map` objects

`redist` provides an extension to `tibble`s, called a `redist_map`. 
This builds on the `sf` tibble class and adds an adjacency column to the data.
The adjacency column is modeled on each state's particular definition of contiguity.
For example, in Washington, contiguity across water is not considered unless there is a ferry route there.
This makes it easy to plug into `redist`'s sampling functions, like [`redist_smc`](https://alarm-redist.org/redist/reference/redist_smc.html) or [`redist_mergesplit`](https://alarm-redist.org/redist/reference/redist_mergesplit.html).

Even for those not running simulation analyses, the `redist_map` can still be useful, as it is just an extension of the `sf` class.
Each object contains the enacted 2010 and 2020 congressional plans.
They also contain census variables such as total population (`pop`), voting age population `vap`, and the major racial categories (e.g. `pop_white` or `vap_black`).

`redist_map`s also include election data, which are retabulations of [VEST's election data](https://dataverse.harvard.edu/dataverse/electionscience) to census shapes.
Election data follows a pattern of `office_year_party_candidate`, like `pre_16_dem_cli` for Clinton's 2016 presidential votes.
Year averages take the form `adv_year` for Democratic votes and `arv_year` for Republican votes (e.g. arv_20 or adv_20).
Finally, the `redist_map` also includes a `ndv` and `nrv` column, which are the average of Democratic and Republican votes in each shape, respectively. 


Given all this, it interfaces nicely with `ggplot2` for making maps.
Here, we can also load the `ggredist` package for some nice election mapping functions and colors.
To make a precinct map, we could run the following:

```{r}
library(ggplot2)
library(ggredist)
map_wa |> 
    ggplot() +
    geom_sf(aes(fill = ndv / (nrv + ndv))) +
    scale_fill_party_c() +
    theme_map()
```

To make a 2020 district map, we can swap out the `geom` call:

```{r}
map_wa |> 
    ggplot() +
    geom_district(aes(group = cd_2020, fill = ndv, denom = nrv + ndv)) +
    scale_fill_party_c() +
    theme_map()
```





# Using `redist_plans` with your own data

Once you've downloaded some plans, we've made it easy for you to add your own data to the `redist` framework.
To demo this, we can look at the NY plans passed last week (February 2024).

First, we can load in the map and plans for New York:

```{r}
map_ny <- alarm_50state_map('NY')
plans_ny <- alarm_50state_plans('NY')
```

Next, we can read in the block assignment data for the commission plan and the legislature plan.
For convenience of the example, we can grab this data from a [GitHub repo](https://github.com/christopherkenny/ny-baf). 

```{r}
# commission link
comm_link <- 'https://raw.githubusercontent.com/christopherkenny/ny-baf/main/data/A09310.csv'

# download the legislature xlsx file
leg_link <- 'https://github.com/christopherkenny/ny-baf/raw/main/data-raw/congressional_plan_equivalency.xlsx'
temp_leg <- tempfile(fileext = '.xlsx')
curl::curl_download(leg_link, temp_leg)

# read in the data
baf_commission <- readr::read_csv(comm_link, col_types = 'ci') |> 
  dplyr::rename(A09310 = district)
baf_legislature <- readxl::read_excel(temp_leg, col_types = c('text', 'text')) |> 
  dplyr::rename(
    GEOID = Block,
    commission2024 = `DistrictID:1`
  ) |> 
  dplyr::mutate(commission2024 = as.integer(commission2024))
```

Finally, `alarmdata` provides a function `alarmdata::alarm_add_plan()` to add the new reference plans to the underlying `redist_plans` object.

```{r}
plans_ny <- plans_ny |>
  alarm_add_plan(ref_plan = baf_commission, map = map_ny, name = 'A09310') |>
  alarm_add_plan(ref_plan = baf_legislature, map = map_ny, name = 'commission2024')
```

This gives a full `redist_plans` object with the new plans added in.
We can do anything we want with this.
For example, stealing a bit of code from [our 50-states website](https://github.com/alarm-redist/alarm-redist.github.io/blob/main/_fifty-states/fifty-states.R), we can look at the Democratic vote shares across districts:

First, we can define some helper functions for the plot:

```{r}
lbl_party <- function(x) {
  dplyr::if_else(x == 0.5, "Even",
    paste0(dplyr::if_else(x < 0.5, "R+", "D+"), scales::number(200 * abs(x - 0.5), 1))
  )
}

r_geom <- function(...) {
  ggplot2::geom_point(
    ggplot2::aes(x = as.integer(.data$.distr_no),
                 y = e_dvs,
                 color = .data$draw, 
                 shape = .data$draw),
    ...
  )
}
```

Then, we can make a simple plot of Democratic vote share by district:

```{r}
library(ggplot2)
library(ggredist)
redist.plot.distr_qtys(plans_ny, e_dvs,
                       color_thresh = 0.5,
                       size = 0.04 - sqrt(8) / 250, alpha = 0.4,
                       ref_geom = r_geom
) +
    geom_hline(yintercept = 0.5, color = '#00000055', size = 0.5) +
    scale_y_continuous('Two-party vote margin', labels = lbl_party) +
    labs(x = 'Simulated districts, ordered by Democratic vote margin') +
    annotate('text',
             x = 3.5, y = sort(plans_ny$e_dvs[1:26])[3],
             label = 'Commission 2024', hjust = -0.05, size = 3.5,
             color = '#52796F'
    ) +
    annotate('text',
             x = 1.5, y = min(plans_ny$e_dvs[27:52]),
             label = 'A09310', hjust = 0.05, size = 3.5,
             color = '#A09310'
    ) +
    annotate('text',
             x = 3.5, y = sort(plans_ny$e_dvs[53:78])[3],
             label = 'Cervas', hjust = -0.05, size = 3.5,
             color = 'black'
    ) + 

    scale_color_manual(values = c('#52796F', '#A09310', 'black', ggredist$partisan[2], ggredist$partisan[14]),
                       labels = c('pt', 'Rep.', 'Dem.'), guide = 'none') +
    scale_shape_manual(values = c(16, 17, 18), guide = 'none') +
    theme_bw()
```

# Cache data across projects

By default, all downloads are directed to a temporary directory.
Each time you reload R, you need to re-download the data.
To avoid this, you can set the `alarm.use_cache` option to a directory where you want to store the data.

```r
options(alarm.use_cache = TRUE)
```

This can be set on a by-session basis, but we recommend setting it in your `.Rprofile` so that it is set every time you start R. To open this file, use the following command:

```r
usethis::edit_r_profile()
```
