---
title: "Fifty States Data Descriptor"
description: |
  A detailed description of the 50-State Redistricting Simulations 
  and new software to help you use them.
author:
    - name: Cory McCartan
      affiliation: Department of Statistics, Harvard University
      location: Cambridge, MA
      email: cmccartan@fas.harvard.edu
      url: https://www.corymccartan.com
    - name: Christopher T. Kenny
      affiliation: Department of Government, Harvard University
      location: Cambridge, MA
      email: christopherkenny@fas.harvard.edu
      url: https://www.christophertkenny.com/
    - name: Tyler Simko
      affiliation: Department of Government, Harvard University
      location: Cambridge, MA
      email: tsimko@g.harvard.edu
      url: https://tylersimko.com/
    - name: George Garcia III
      affiliation: Department of Economics, Massachusetts Institute of Technology
      location: Cambridge, MA
      email: grg3@mit.edu
    - name: Kevin Wang
      affiliation: Harvard College
      location: Cambridge, MA
      email: kpwang@college.harvard.edu
      url: https://scholar.harvard.edu/kevinwang
    - name: Melissa Wu
      affiliation: Harvard College
      location: Cambridge, MA
      email: mzwu@college.harvard.edu
    - name: Shiro Kuriwaki
      affiliation: Department of Political Science, Yale University
      location: New Haven, CT
      email: shiro.kuriwaki@yale.edu
      url: https://www.shirokuriwaki.com/
    - name: Kosuke Imai
      affiliation: Departments of Government and Statistics, Harvard University
      location: Cambridge, MA
      email: imai@harvard.edu
      url: https://imai.fas.harvard.edu/
date: 2022-07-28
---


```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = FALSE)
```


__It's been a long redistricting year.__
We've been tracking passed maps while conducting simulations in the 44 states with congressional districts.
We are now finalizing some re-runs of states with new validation steps based on additional diagnostics, to ensure a high quality and accurate data product.
So, we've written up a more detailed draft of what we did, how we did it, and how we checked our work.
Most importantly, it introduces some tools so that you can use the data we've generated.
It's all open source and the redistricting plans generated are in the public domain.

Read the detailed description of the our process and the data: [Simulated redistricting plans for the analysis and evaluation of redistricting in the United States: 50stateSimulations](https://arxiv.org/abs/2206.10763). The abstract is listed below.

> This article introduces the 50stateSimulations, a collection of simulated congressional districting plans and underlying code developed by the Algorithm-Assisted Redistricting Methodology (ALARM) Project. The 50stateSimulations allow for the evaluation of enacted and other congressional redistricting plans in the United States. While the use of redistricting simulation algorithms has become standard in academic research and court cases, any simulation analysis requires non-trivial efforts to combine multiple data sets, identify state-specific redistricting criteria, implement complex simulation algorithms, and summarize and visualize simulation outputs. We have developed a complete workflow that facilitates this entire process of simulation-based redistricting analysis for the congressional districts of all 50 states. The resulting 50stateSimulations include ensembles of simulated 2020 congressional redistricting plans and necessary replication data. We also provide the underlying code, which serves as a template for customized analyses. All data and code are free and publicly available. This article details the design, creation, and validation of the data.

To help make things more usable for those who don't simulate redistricting plans in their free time, we are also excited to (soft) launch a new R package, [`alarmdata`](https://github.com/alarm-redist/alarmdata).
This package provides a simplified interface to download the underlying geographic data, generated plans, all sorts of summary statistics, and state-by-state documentation.

The package can be installed with:

```{r}
#| eval: false
#| echo: true
remotes::install_github('alarm-redist/alarmdata')
```


Thank you to the Harvard Data Science Initiative and Microsoft for computational support.
